Strong lensing selection effects

### 2023-05-31 paper/seleff.tex
minmu

Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
\begin{enumerate}
\item The lens must produce multiple images of the source
\item The multiple images must be detected and resolved
\item The system must be recognized as a lens by the survey.
\end{enumerate}
Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
This condition excludes objects with a very diffuse mass distribution from the population of lenses.
Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples. Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.

@@

The cumulative effect of these conditions can be summarised as follows.
%In general, 
The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
\begin{equation}\label{eq:one}
\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\psel(\psilens,\psisource|S).
\end{equation}
In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.

@@

>>>>

{\bf
Strong lenses, however, are a biased subset of the general population of galaxies and background sources.
In order for a lens-source pair to be included in a strong lensing survey, the following conditions must be met:
\begin{enumerate}
\item The lens must produce multiple images of the source
\item The multiple images must be detected and resolved
\item The system must be recognized as a lens by the survey.
\end{enumerate}
Condition 1 poses constraints on the mass distribution of the lens, as well as on the geometry of the system. 
First of all, not all lenses are capable of producing multiple images, but only those whose surface mass density $\Sigma(\boldsymbol\theta)$ is larger than the critical surface mass density for lensing $\Sigma_{\mathrm{cr}}$ at at least one position $\boldsymbol\theta$ \citep{SEF92}.
This condition excludes objects with a very diffuse mass distribution from the population of lenses.
Second, objects with a higher concentration of mass allow for a larger portion of the source plane to be multiply imaged, and therefore are over-represented in lens samples. Third, at fixed lens properties, sources that are located further away along the line of sight are more likely to be lensed.

@@

The cumulative effect of these conditions can be summarised as follows.
}
%In general, 
The probability distribution $\prsl$ of a sample of strong lenses from a given survey with selection criterion $S$ is given by \citep{Son22}:
\begin{equation}\label{eq:one}
\prsl(\psilens,\psisource|S) \propto \prlens(\psilens)\prsource(\psisource)\psel(\psilens,\psisource|S).
\end{equation}
In the equation above, $\psilens$ is the set of parameters describing the properties of foreground galaxies that are relevant for lensing, such as their redshift and mass distribution; $\psisource$ is the set of parameters describing background sources; $\prlens$ and $\prsource$ describe the parent distribution of foreground galaxies and background sources in the absence of lensing; and $\psel$ describes the probability of selecting a lens-source system with parameters $\psilens$ and $\psisource$ given the criterion $S$ used to define a lens.
This last factor takes into account both physical effects, that is whether a lens with parameters $\psilens$ produces a strongly lensed image of a source with parameters $\psisource$, and survey selection effects, that is whether such an image can be detected and recognised as a strong lens.

@@

\subsection{Alternative lens definitions}

{\bf
Our extended source analysis was based on a peak detection criterion to define strong lensing events, as explained in section \ref{ssec:lensdefext}.
However, many previous studies aimed at simulating populations of lenses adopted a magnification-based definition, in which a lens-source system is classified as a strong lens if the magnification of the source is larger than a minimum value \citep[see e.g.][]{Col15, Rob++20}.
In order to check whether our results are robust to the choice of definition of a strong lensing event, we re-analysed the fiducial sample with this alternative definition. %repeated the analysis of the fiducial sample with extended sources. 
In particular, we defined as strong lenses all systems with a detected background source and $\mu_{\mathrm{tot}} > 3$, where $\mu_{\mathrm{tot}}$ is the total magnification (obtained by integrating the flux of all of the lensed images and dividing it by the intrinsic source flux).
The total number of strong lenses increases by $34\%$ with this new definition, mostly due to the inclusion of fainter sources with no detected multiple images.
\Fref{fig:minmubias} shows the strong lensing bias on a few selected quantities (stellar population synthesis mismatch parameter, projected dark matter mass within $5$~kpc, source magnitude and source half-light radius).

}

%However, many previous simulation studies adopted criteria based on magnification \citep[see e.g.][]{Col15, Rob++20}.

However, in the literature it is common to find alternative criteria based on magnification \citep[see e.g.][]{Col15,

it is common to find 
In the 
Other simulation works, however, use a

Our analysis was based on the definition of a strong lens given in sections \ref{ssec:lensdefpoint} and \ref{ssec:lensdefext}. 
These are not the only possible definitions.
In the case of an extended source, in particular, it is common to find in the literature magnification-based
